The four postulates of quantum mechanics are three

### 2020-11-07 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex

The tensor product postulate does not appear in all axiomatizations of
quantum mechanics: it has even been called ``postulate 0'' in some
literature \cite{zurek}. A widespread belief is that it is a direct
consequence of the superposition principle, and it is hence not a
necessary axiom. {\em This belief is mistaken}: the superposition
principle is encoded into the quantum axioms by requiring that the
state space is a {\em linear} vector space. This is, by itself,
insufficient to single out the tensor product, as other linear
products of linear spaces exist, such as the direct product, the
topological product or the direct sum of vector spaces, which are used
in classical mechanics to combine state spaces of linear systems.
This belief may have arisen from the seminal book of Dirac
\cite{diracbook}, who introduces tensor products (Chap.~20) by
appealing to linearity. However, he adds the seemingly innocuous
request that the product among spaces be distributive (rather,
bilinear), which is equivalent to postulating tensor products (or
linear functions of them). This is not an innocuous request. For
example it does not hold where the composite vector space of two
linear spaces is described by the direct product, e.g.~in classical
mechanics, for two strings of a guitar: it is not distributive.
[General classical systems, not only linear ones, are also composed
through the direct product.] Of course, Dirac is not constructing an
axiomatic formulation, so his `sleight of hand' can be forgiven. In
contrast, von Neumann (\cite{vonneumannbook} Chap.~VI.2, also
\cite{jauch}) introduces tensor products by noticing that this is a
natural choice in the position representation of wave mechanics (where
they were introduced in \cite{weyl,epr}), and then {\em explicitly
  postulates} them in general: ``This rule of transformation is
correct in any case for the coordinate and momentum operators [...]
and it conforms with the [observable axiom and its linearity
principles], we therefore postulate them generally.''
\cite{vonneumannbook}.  More mathematical or conceptually-oriented
modern formulations (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang})
introduce this postulate explicitly.  An interesting alternative is
provided in \cite{ballentinebook,ballentinepaper}: after introducing
tensor products, Ballentine verifies a posteriori that they give the
correct laws of composition of probabilities. Similarly, Peres uses
relativistic locality \cite{peres}. While these procedures seemingly
bypass the need to postulate the tensor product, they do not guarantee
that this is the {\em only} possible way of introducing composite
systems in quantum mechanics. In the framework of quantum logic,
tensor products arise from some additional conditions \cite{matolcsi}
which (in contrast to what is done here) are not connected to the
other postulates. Similar approaches appear in \cite{marmo,aerts}
where tensor products were obtained by specifying additional physical
or mathematical requirements. In quantum field theory one tends to
avoid problems connected with tensor products of infinite dimensional
spaces by focusing on algebraic commutation structures,
e.g.~\cite{giddins,roos}.  In particular, the recent MIP*=RE result
\cite{mipre} implies that, in infinite dimensions, the tensor product
is strictly less computationally powerful than the commutation
structures, emphasizing the difference among these two structures, at
least for the infinite-dimensional case. We will consider the
non-relativistic setting here.

In this paper we derive the tensor product postulate (which, hence,
loses its status of postulate) from two other postulates of quantum
mechanics: the state postulate and the measurement postulate. We start
from the natural definition of a composite system as the set of two
(or more) quantum systems, in the sense that the composite system is
made of system $A$ {\em and} (joined with) system $B$, namely a set
whose elements are the two systems and nothing else. We will focus on
kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
other restrictions to the state space are present: it is possible to
prepare each subsystem of a composite system in a state that is
independent of the other systems (preparation independence).  This is
the only case in which the tensor product can be properly employed
\cite{susskind,zanardi,zanardilloyd}: the Hilbert space of composite
systems that have restrictions is {\em not} the tensor product of the
component spaces, but a subspace of it (e.g.~the anti-symmetric
subspace for fermions). Typically this is ignored in the literature,
since the tensor product formalism is very convenient and is often
used also in these cases. Moreover, in accordance with the measurement
postulate, we require that the probability distribution of the
measurement outcomes of one system is independent of the other systems' (statistical
independence). As detailed below, statistical independence is not an
additional requirement: it is already contained in the measurement
postulate.

From the above definition of composite system, it follows that there
must be a map $M$ that connects the states of the subsystems to the
states of the composite system. A quantum state is a ray in Hilbert
space, namely a set of vectors. The map $M$ on states corresponds to a
map $m$ on vectors. We then show that preparation independence and
statistical independence imply three conditions on the map $m$:
(H1)~totality: the map is defined on all states of the subsystems;
(H2)~bilinearity: the map is bilinear thanks to the fundamental
theorem of projective geometry; (H3)~span surjectivity: the span of
the image of map coincides with the full composite Hilbert space.  We
then prove that, if these three conditions H1, H2 and H3 hold, then
the map $m$ is the tensor product, namely the Hilbert space of the
composite system is a tensor product of the components: the tensor
product ``postulate'', which hence loses its status of a postulate. An
overview of all these logical implications is given in
Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
argument. Refer to the Supplementary Information for a more rigorous
proof of the \emph{same} argument.

@@

We use the axiomatization of quantum mechanics based on the following
postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The state of a
system is described by a ray $\pj{\psi}$ corresponding to a set of
non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
system's observable properties are described by self-adjoint operators
acting on that space; (b)~The probability that a measurement of a
property $X$, described by the operator with spectral decomposition
$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
$x$ given that the system is in state $\pj{\psi}$ is
$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
space of a composite system is given by the tensor product of the
spaces of the component systems; (d)~The time evolution of an isolated
system is described by a unitary operator acting on a vector
representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
equivalently, by the Schr\"odinger equation. The rest of quantum
theory can be derived from these axioms. While some axiomatizations
introduce further postulates, we will be using only (a) and (b) to
derive (c), so the above are sufficient for our aims.

@@

The definition of a composite system as containing {\em only} the
collection of the subsystems means that any preparation of both
subsystems independently must correspond to the preparation of the
composite system. Since states are defined by postulate (a) as rays in
the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
projective spaces and the Cartesian product is the set of all possible
pairs) and returns a state in the projective space $\pj{\cal C}$ for
the composite. The map $M$ that acts on rays corresponds to a map
$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
Hilbert spaces $\cal A$, $\cal B$ and $\cal C$.  Namely,
$\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign indicates the
elements in the projective space. We will prove that the map $m$ is
the tensor product. We focus on pure states here: the argument can be
extended to mixed states using standard tools \cite{nielsenchuang}.

@@

Postulate (b) contains the connection between quantum mechanics
and probability theory. It must then implicitly contain the
axiomatization of probability, e.g.~see
\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
the joint probability of events $a$ and $b$ given $z$ is
$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Then the events $a$ and $b$ are
independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$ (this
is the {\em definition} of independent events). Since the Born
probability formula in postulate (b) contains only quantities of the
system (the state and the observable's eigenstates {\em of the
  system}), it implies the ``statistical independence'' between
different systems if they are prepared independently. Namely, the
outcome probability of one is independent of any properties of any
other in the absence of any dynamical coupling and supposing
independent preparation. [Of course, the dynamics may couple the state
of the system to other systems so that the outcomes may depend on what
happens to other systems.  Analogously, one may prepare the system in
a way that depends on other systems. But we consider independent
preparations and do not consider dynamics, as it is sufficient for our
aims.] We can then formalize ``statistical independence'' of
independently prepared systems by saying that their Born rule
satisfies:
\begin{eqnarray}
  p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
    B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
    A})\:p(b|\psi_{\cal B})
\labell{bindip}\;,
\end{eqnarray}
where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
B$ respectively. The first equality in equation \eqref{bindip} embodies the
definition of statistical independence, the second follows from the
form of the Born rule for independently prepared systems (the
probability depends only on the state and observables of each system
on its own).

@@

Since the square of the inner product is preserved, orthogonality and
the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
colinear transformation by definition. In this case, the fundamental
theorem of projective geometry \cite{fun} applies, which tells us that
a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
Moreover, conservation of probability further constrains it to be
either linear or antilinear. This tells us that the corresponding $m$
is either linear or antilinear in the first argument. Namely, if equation
\eqref{questa} holds, then
\begin{align}
\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
 \\\mbox{ or }
\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
\end{align}
We can ignore the antilinear case \eqref{h2b}, which simply
corresponds to the map $m$ mapping kets of one or both subsystems into
bras before outputting the composite system. All these maps are
trivially physically equivalent, since the dual-space description of a
quantum system through bras is equivalent to the description using
kets, or in other words, the wavefunction and its complex-conjugate
contain identical information (for example, in his first papers
Schr\"odinger used both signs in his equation: effectively writing
{\em two} equivalent equations with complex-conjugate solutions
\cite{sch}). We can now repeat the same analysis for the second
argument of $m$ to conclude that it is a bilinear map, condition (H2).

@@

Step 2: use the universal property. The tensor product is uniquely
characterized, up to isomorphism, by a universal property regarding
bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
through $T$ uniquely.  This means that there exists a {\em unique}
$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
words, the following diagram commutes:
\begin{center}
\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
& \mathcal{C}
\end{tikzcd}
\end{center}
Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
a bilinear operator (property H2), thanks to the universal property of
the tensor product we can find a unique linear operator $\hat{m} :
\mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $m(a, b) =
\hat m(a \otimes b)$. The set $\{ \hat m(a_i\otimes b_j)$ with
$|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and ${\cal B}\}$
forms a basis for $\cal C$, since $\hat m(a_i\otimes b_j)=m(a_i,b_j)$
and we have shown above that the latter is a basis.  Thus,
\begin{align} 
  &\<\hat m(a_i\otimes b_j)|\hat m(a_k\otimes b_\ell)
\>_{\mathcal{C}}=\<m(a_i,
  b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
  \delta_{ik}\delta_{j\ell}
 = \<a_i\otimes
  b_j| a_k \otimes b_\ell\>_{\otimes},
\labell{ecco}\; 
\end{align}
where we used the orthonormality of the bases and the fact that
$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal
  A}\otimes{\cal B}$. The function $\hat{m}$, then, is an isomorphism
over all elements of the basis $|a_i\otimes b_j\>$. It is then an
isomorphism between $\cal C$ and ${\cal A}\otimes{\cal B}$. Namely,
$|m(a_i,b_j)\>_{\cal C} = |\hat{m}(a_i\otimes b_j)\>_{\cal C} \cong
|a_i\otimes b_j\>_\otimes$, where the first two vectors (in $\cal C$)
are isomorphic to the last vector (in the tensor product space ${\cal
  A}\otimes{\cal B}$). Namely, $\hat{m}$ maps bijectively to the
tensor product, which implies the isomorphism ${\cal C}\cong{\cal
  A}\otimes{\cal B}$, i.e.~their equivalence through the map
$\hat{m}$. Moreover, since ${\cal C}\cong{\cal A}\otimes{\cal B}$,
then the map $m : \cal A \times \cal B \to \cal C$ is equivalent to
the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$,
i.e.~the maps $m$ and $\otimes$ are also equivalent through $\hat
m$.$\square$

>>>>

The tensor product postulate does not appear in all axiomatizations of
quantum mechanics: it has even been called ``postulate 0'' in some
literature \cite{zurek}. A widespread belief is that it is a direct
consequence of the superposition principle, and it is hence not a
necessary axiom. {\em This belief is mistaken}: the superposition
principle is encoded into the quantum axioms by requiring that the
state space is a {\em linear} vector space. This is, by itself,
insufficient to single out the tensor product, as other linear
products of linear spaces exist, such as the direct product, the exterior/wedge roduct or the direct sum of vector spaces, which is used
	in classical mechanics to combine state spaces of linear systems. These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \otimes c$ where $0$ is the zero vector. Also, in the tensor product $r * (a \otimes b) = (r*a) \otimes b = a \otimes (r*b)$ while in the direct product $r * (a \times b) = r*a \times r*b$.}

	This belief may have arisen from the seminal book of Dirac
	\cite{diracbook}, who introduces tensor products (Chap.~20) by
	appealing to linearity. However, he adds the seemingly innocuous
	request that the product among spaces be distributive (rather,
	bilinear), which is equivalent to postulating tensor products (or
	linear functions of them). This is not an innocuous request. For
	example it does not hold where the composite vector space of two
	linear spaces is described by the direct product, e.g.~in classical
	mechanics, for two strings of a guitar: it is not distributive.
	[General classical systems, not only linear ones, are also composed
	through the direct product.] Of course, Dirac is not constructing an
	axiomatic formulation, so his `sleight of hand' can be forgiven. In
	contrast, von Neumann (\cite{vonneumannbook} Chap.~VI.2, also
	\cite{jauch}) introduces tensor products by noticing that this is a
	natural choice in the position representation of wave mechanics (where
	they were introduced in \cite{weyl,epr}), and then {\em explicitly
		postulates} them in general: ``This rule of transformation is
	correct in any case for the coordinate and momentum operators [...]
	and it conforms with the [observable axiom and its linearity
	principles], we therefore postulate them generally.''
	\cite{vonneumannbook}.  More mathematical or conceptually-oriented
	modern formulations (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang})
	introduce this postulate explicitly.  An interesting alternative is
	provided in \cite{ballentinebook,ballentinepaper}: after introducing
	tensor products, Ballentine verifies a posteriori that they give the
	correct laws of composition of probabilities. Similarly, Peres uses
	relativistic locality \cite{peres}. While these procedures seemingly
	bypass the need to postulate the tensor product, they do not guarantee
	that this is the {\em only} possible way of introducing composite
	systems in quantum mechanics. In the framework of quantum logic,
	tensor products arise from some additional conditions \cite{matolcsi}
	which (in contrast to what is done here) are not connected to the
	other postulates. Similar approaches appear in \cite{marmo,aerts}
	where tensor products were obtained by specifying additional physical
	or mathematical requirements. In quantum field theory one tends to
	avoid problems connected with tensor products of infinite dimensional
	spaces by focusing on algebraic commutation structures,
	e.g.~\cite{giddins,roos}.  In particular, the recent MIP*=RE result
	\cite{mipre} implies that, in infinite dimensions, the tensor product
	is strictly less computationally powerful than the commutation
	structures, emphasizing the difference among these two structures, at
	least for the infinite-dimensional case. We will consider the
	non-relativistic setting here.

	In this paper we derive the tensor product postulate (which, hence,
	loses its status of postulate) from two other postulates of quantum
	mechanics: the state postulate and the measurement postulate. We start
	from the natural definition of a composite system as the set of two
	(or more) quantum systems. The composite system is therefore
	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else},
	it's state being fully described by all observables and only the observables
	that fully describe its constituents. The first key insight is that first two postulates already assume that, at least in some cases, the probability distribution of the
	measurement outcomes of one system is independent of the other systems' (statistical independence). While we will see specifically how the Born implies this, it should be clear conceptually that we would not be able to talk about a system in the first place if we would not be able to characterize it independently. The second key insight is that, since all measurement
	outcomes and all states of the subsystems plus the composite live in the same probability space, we will automatically have a map $M$ that for a pair of subsystems' states gives us the composite state that corresponds to the statistically independent case. These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states). So the work consists of two interrelated efforts: a physical argument that start from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.

	We will mainly focus on
	kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
	other restrictions to the state space are present: it is possible to
	prepare each subsystem of a composite system in a state that is
	independent of the other systems (preparation independence).  This is
	the only case in which the tensor product can be properly employed
	\cite{susskind,zanardi,zanardilloyd}: the Hilbert space of composite
	systems that have restrictions is {\em not} the tensor product of the
	component spaces, but a subspace of it (e.g.~the anti-symmetric
	subspace for fermions). Typically this is ignored in the literature,
	since the tensor product formalism is very convenient and is often
	used also in these cases, but it is something we need to bring out.

	While conceptually the derivation is straight forward, it requires the
	use of projective space, which may be unfamiliar to the reader. This is
	necessary because events in probability space do not correspond to vectors in the Hilbert space, but to subspaces. If $\psi$ is a vector, the state corresponds to the ray $\pj{\psi}$, which is the subspace spanned by $\psi$. Note that ``ray'' can be misleading since $\pj{\psi}$ is really a complex { \em plane }. In the same way, the event $X = x_0$ corresponds to the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$. The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projectives spaces. The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.

	This is done by showing that preparation independence and
	statistical independence imply three conditions on the map $m$:
	(H1)~totality: the map is defined on all states of the subsystems;
	(H2)~bilinearity: the map is bilinear thanks to the fundamental
	theorem of projective geometry; (H3)~span surjectivity: the span of
	the image of map coincides with the full composite Hilbert space.  We
	then prove that, if these three conditions H1, H2 and H3 hold, then
	the map $m$ is the tensor product, namely the Hilbert space of the
	composite system is a tensor product of the components: the tensor
	product ``postulate'', which hence loses its status of a postulate. An
	overview of all these logical implications is given in
	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
	argument. Refer to the Supplementary Information for a more rigorous
	proof of the \emph{same} argument.

@@

	We start with the axiomatization of quantum mechanics based on the following
	postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state of a
	system is described by a ray $\pj{\psi}$ corresponding to a set of
	non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
	system's observable properties are described by self-adjoint operators
	acting on that space; (b)~The probability that a measurement of a
	property $X$, described by the operator with spectral decomposition
	$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
	$x$ given that the system is in state $\pj{\psi}$ is
	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
	space of a composite system is given by the tensor product of the
	spaces of the component systems; (d)~The time evolution of an isolated
	system is described by a unitary operator acting on a vector
	representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
	equivalently, by the Schr\"odinger equation. The rest of quantum
	theory can be derived from these axioms. While some axiomatizations
	introduce further postulates, we will be using only (a) and (b) to
	derive (c), so the above are sufficient for our aims.

@@

	The definition of a composite system as containing {\em only} the
	collection of the subsystems means that any preparation of both
	subsystems independently must correspond to the preparation of the
	composite system. Since states are defined by postulate (a) as rays in
	the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
	\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
	the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
	projective spaces and the Cartesian product is the set of all possible
	pairs) and returns a state in the projective space $\pj{\cal C}$ for
	the composite. To visualize this geometrically, given a ray (a complex plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex plane) in $\cal C$.  Our final goal will be to find a map
	$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
	Hilbert spaces $\cal A$, $\cal B$ and $\cal C$ consistently with $M$. Namely,
	$\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign indicates the
	elements in the projective space. Again geometrically, $m$ takes a vector in each of $\cal A$ and $\cal B$, and returns a vector in $\cal C$ and we want this to be consistent with $M$ such that vectors picked from the same rays will return vectors in the same ray. We will prove that the map $m$ is
	the tensor product. We focus on pure states here: the argument can be
	extended to mixed states using standard tools \cite{nielsenchuang}.

@@

	Postulate (b) contains the connection between quantum mechanics
	and probability theory. It must then implicitly contain the
	axiomatization of probability, e.g.~see
	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
	the joint probability of events $a$ and $b$ given $z$ is
	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. This essentially the definition of conditional probability: $p(b|z\wedge a) = \frac{p(a\wedge b|z)}{p(a|z)}$. Two events $a$ and $b$ are defined to be 
	independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$. However, it is more instructive to define independence with $p(b|z\wedge a)=p(b|z)$. This makes it more evident that the probability of $b$ does not depend on conditioning on $a$. If we imagine $a$ and $b$ to be outcomes for particular measurements, we are saying that the probability of observing $b$ is not a function of the value measured by $a$. We can recover the first definition from the second. In fact, $p(b|z\wedge a)=\frac{p(a\wedge b|z)}{p(a|z)} =p(b|z)$ which leads to $p(a\wedge b|z)=p(a|z)\:p(b|z)$.

	We note that the Born
	probability formula in postulate (b) is of the form $p(a|\psi_{\cal A}$ and contains only quantities of the
	system (the pure state $\psi_{\cal A}$ and the observable's eigenstates $a$). Now, suppose that, unknown to us, an additional second system was prepared in a pure state $\psi_{\cal B}$ and consider $p(a|\psi_{\cal A}\wedge \psi_{\cal B})$. In what cases are allowed to use the Born rune? That is, in what cases we can say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$? Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$ are statistically independent. In this sense, postulate (b) already implies  ``statistical independence'' between
	different systems if they are prepared independently. Namely, the
	outcome probability of one is independent of any properties of any
	other in the absence of any dynamical coupling and supposing
	independent preparation. [Of course, the dynamics may couple the state
	of the system to other systems so that the outcomes may depend on what
	happens to other systems.  Analogously, one may prepare the system in
	a way that depends on other systems. But we consider independent
	preparations and do not consider dynamics, as it is sufficient for our
	aims.] We can then formalize ``statistical independence'' of
	independently prepared systems by saying that their Born rule
	satisfies:
	\begin{eqnarray}
	p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
		B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
		A})\:p(b|\psi_{\cal B})
	\labell{bindip}\;,
	\end{eqnarray}
	where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
	preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
	$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
	B$ respectively. The first equality in equation \eqref{bindip} embodies the
	definition of statistical independence, the second follows from the
	form of the Born rule for independently prepared systems (the
	probability depends only on the state and observables of each system
	on its own).

@@

	Since the square of the inner product is preserved, orthogonality and
	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
	colinear transformation by definition. Geometrically, recall that $M_b$ maps rays to rays. The fact that $M_b$ is colinear means that it also maps higher order subspaces to higher order subspaces (lines to lines, planes to planes, and so on) while preserving inclusion (if a line is within a plane, the mapped line will be within the mapped plane). In this case, the fundamental
	theorem of projective geometry \cite{fun} applies, which tells us that
	a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
	Moreover, conservation of probability further constrains it to be
	either linear or antilinear. This tells us that the corresponding $m$
	is either linear or antilinear in the first argument. Namely, if equation
	\eqref{questa} holds, then
	\begin{align}
	\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
	\\\mbox{ or }
	\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
	\end{align}
	In this setting, the antilinear case \eqref{h2b} corresponds to a change of convention (much like a change of sign in the symplectic form for classical mechanics) and can be ignored. Given a Hilbert space, in fact, we can imagine replacing all wavefunctions and all the operators with their complex-conjugate. These changes would effectively cancel out leaving the physics being represented unchanged  (for example, in his first papers
	Schr\"odinger used both signs in his equation: effectively writing
	{\em two} equivalent equations with complex-conjugate solutions
	\cite{sch}). Another way to see this is noting that, given any equation 
	in terms of operators acting on ket vectors, one can write an
	equivalent equation in terms of bra vectors by simply taking the
	Hermitian conjugate of both sides of the first equation. Namely, the
	equation A|psi>=B|phi> is equivalent to $<psi|A^dag=<phi|B^dag$. Mathematically, there is nothing preventing one to represent state preparation and kets to represent measurements. It is a matter of convention. We can now repeat the same analysis for the second
	argument of $m$ to conclude that it is a bilinear map, condition (H2).

@@

	Step 2: use the universal property. The tensor product is uniquely
	characterized, up to isomorphism, by a universal property regarding
	bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
	tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
	than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
	through $T$ uniquely.  This means that there exists a {\em unique}
	$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
	words, the following diagram commutes:
	\begin{center}
		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
			& \mathcal{C}
		\end{tikzcd}
	\end{center}
	Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
	a bilinear operator (property H2), thanks to the universal property of
	the tensor product we can find a unique linear operator $\hat{m} :
	\mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $m(a, b) =
	\hat m(a \otimes b)$. The set $\{ \hat m(a_i\otimes b_j)$ with
	$|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and ${\cal B}\}$
	forms a basis for $\cal C$, since $\hat m(a_i\otimes b_j)=m(a_i,b_j)$
	and we have shown above that the latter is a basis.  Thus,
	\begin{align} 
	&\<\hat m(a_i\otimes b_j)|\hat m(a_k\otimes b_\ell)
	\>_{\mathcal{C}}=\<m(a_i,
	b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
	\delta_{ik}\delta_{j\ell}
	= \<a_i\otimes
	b_j| a_k \otimes b_\ell\>_{\otimes},
	\labell{ecco}\; 
	\end{align}
	where we used the orthonormality of the bases and the fact that
	$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal A}\otimes{\cal B}$. Since the function $\hat{m}$ is a linear function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$ to an orthonormal basis of $\cal C$, $\hat{m}$ is a an isomorphism (a bijection that preserves the mathematical structure) between ${\cal A}\otimes{\cal B}$ and $\cal C$. As ${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert spaces, they are mathematically equivalent and we can directly use the tensor product to represent the composite state space. This means that the map $m : \cal A \times \cal B \to \cal C$ is equivalent to the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the sense that $m \circ \hat m^{-1} = \otimes$.$\square$
